Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-09 paper.qmd
Checked 2D notation

- Make sure langevitour is explained before using it
- Too much code needs to be streamlined
- Remove all the extraneous commented text
- Check how \kD{} and \gD{} are used, and make sure it's sensible

@@

## Assessing the best fit, and hence most accurate 2D layout

@@

When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PACMAP, when the data is structure is higher than two dimensional. A simple example of this can be seen with data simulated to contain five \fD{} Gaussian clusters. Each cluster is essentially a ball in \fD{}, so there is no \2D{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out \fD{}.

@@

### Low density results in a squishing of the 2D representation in all methods {#sec-effect-dens}

Differences in density can arise by sampling at different rates in different subspaces of \pD{}. For example, the data shown in @fig-one-dens_clust-error all lies on a \2D{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely. It was simulated to illustrate the effect of the density difference on layout generated by an NLDR, illustrated using the tSNE results.

>>>>

- Make sure langevitour is explained before using it
- Too much code needs to be streamlined
- Remove all the extraneous commented text (Done)
- Check how \kD{} and \gD{} are used, and make sure it's sensible

@@

## Assessing the best fit, and hence most accurate \gD{} layout

@@

When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PACMAP, when the data is structure is higher than two dimensional. A simple example of this can be seen with data simulated to contain five \fD{} Gaussian clusters. Each cluster is essentially a ball in \fD{}, so there is no \gD{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out \fD{}.

@@

### Low density results in a squishing of the \gD{} representation in all methods {#sec-effect-dens}

Differences in density can arise by sampling at different rates in different subspaces of \pD{}. For example, the data shown in @fig-one-dens_clust-error all lies on a \gD{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely. It was simulated to illustrate the effect of the density difference on layout generated by an NLDR, illustrated using the tSNE results.
